### 2014-08-04 tex/lib/DiagramAAM.tex
initial

\begin{center}
\begin{tabular*}{0.66\textwidth}{@{\extracolsep{\fill}} c c}
  \begin{tikzpicture}[ampersand replacement=\&]
    \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
    { \DSS                                 \& \DSS                                 \\
      \node [lightgray] (m-2-1) {\DSSAbs}; \& \node [lightgray] (m-2-2) {\DSSAbs}; \\
    } ;
    \path[-stealth]
    (m-1-1) edge [lightgray,bend right=40] node [lightgray,left] {$\alpha$} 
                                           node [lightgray,right] {\tiny(2.a)} (m-2-1)
    (m-2-1) edge [lightgray,bend right=40] node [lightgray,right] {$\gamma$} (m-1-1)

    (m-1-2) edge [lightgray,bend right=40] node [lightgray,left] {$\alpha$} (m-2-2)
    (m-2-2) edge [lightgray,bend right=40] node [lightgray,right] {$\gamma$} (m-1-2)

    (m-1-1) edge node [above] {$\Dstep$} (m-1-2)
    (m-2-1) edge node [above] {$\DstepAbs$} 
                 node [below] {\tiny(1.b)} (m-2-2)
    ;
    \draw 
      (m-2-1) node [lightgray,left=0.25cm] {\tiny(1.a)} 
    ;
  \end{tikzpicture}
  &
  \begin{tikzpicture}
    \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
    { \DstepAbs \\
      \Dstep    \\
    } ;
    \path
    (m-1-1) edge node [right] {$\sqsubseteq$} 
                 node [left] {\tiny(2.b)} (m-2-1)
    ;
  \end{tikzpicture}
\end{tabular*}
\end{center}