### 2014-08-04 tex/sections.old/00Abstract.tex

The design and implementation of static analyses is a difficult process.
%
Verifying the correctness of an implementation is often the most painful step.
%
We present Monadic AAM--an extension of the Abstracting Abstract Machines
methodology introduced by Van Horn and Might~\cite{van-horn:2010:aam}--which
captures a large class of both automatically derivable and correct by
construction abstract interpreters.
%
Monadic AAM is part methodology and part toolkit.
%
In the methodology, semantics are designed in a monadic extension of
traditional AAM.
%
Once designed, a large class of known analyses can be automatically recovered
using our language agnostic toolkit.
%
Both the computational and correctness properties of our framework are realized
through a restricted class of monad transformers.
%
Consequently our framework enjoys the benefits of being highly compositional
and demands minimal proof burdens from the user.