### 2014-08-04 tex/sections.old/01Introduction.tex

\subsection{Basic Order Theory}

\noindent [TODO: Partial Orders. Define ``discrete ordering''.]

\noindent [TODO: Products.]

\noindent [TODO: Function Spaces.]

\noindent [TODO: Sets.]

\noindent [TODO: Monotonicity.]

\noindent [TODO: Fixpoints.]

\subsection{Basic Abstract Interpretation Concepts}

\noindent [TODO: Small-step semantics.]

\noindent [TODO: Galois Connections.]

\noindent [TODO: An extremely simple example (borrow this from Cousot).]

\noindent [TODO: A calculational derivation of the previous example.]

\subsection{Small Step Semantics}

\noindent [TODO: Lambda Calculus Reduction Semantics.]

\noindent [TODO: Lambda Calculus Small Step Semantics.]

\noindent [TODO: CPS motivation and examples.]