### 2014-08-04 tex/sections.old/05Modularity.tex

\label{section:modularity}

The Abstracting Abstract Interpreters (AAI) framework developed by Might and
Van Horne provides an axis on which to develop a spectrum of analyses of
varying precision and intensional behavior.
%
The key benefit of their approach is that it generalizes to arbitrary languages
and their state spaces.

This framework provides a separate orthogonal axis on which extensions to
abstract interpretation can be explained, and is similarly modular and language
agnostic.
%
In particular, the techniques of AAI and Monadic Abstract Interpretation can be
combined arbitrarily.

%%%%%%%%%%%%%%%%%%%% Related Work %%%%%%%%%%%%%%%%%%%%