### 2014-08-04 tex/sections/03AAMByExample/02AbstractStateSpace/00Cut.tex

In the AAM methodology, we first identify recursion in the state space and
replace it with indirection through addresses and a store.
%
\begin{figure}[H]
\haskell{sections/03AAMByExample/02AbstractStateSpace/00Cut.hs}
\caption*{Cutting Recursion}
\end{figure} 
\noindent
%
Abstract time is introduced into the state space as a constantly increasing
value from which to allocate fresh addresses.

%--%

As a result of this change in the state space, the semantics will need to
change accordingly.
%
Variable lookup must fetch an address from the environment, and then fetch the
corresponding value bound to that address from the store.
%
Variable binding must allocate globally fresh addresses from the current time,
binding the formal parameter to the address in the environment and the address
to the value in the store.