### 2014-08-04 tex/sections/03AAMByExample/02AbstractStateSpace/01AAM.tex

Now that recursion is eliminated from the state space we take aim at unbounded
parts of the state space, with the addresses and time that we just introduced
being the first to go.
%
Rather than make a single choice of abstract address and time, we leave this
choice as a parameter.
%
All that is required is:
\begin{itemize}
\item a time \h|tzero| to begin execution
\item a function \h|tick| which moves time forward
\item a function \h|alloc| which allocates an address for binding a formal
      parameter at the current time
\end{itemize}
%
We encode this interface using a Haskell type class \h|AAM| with associated
types \h|Addr| and \h|Time|.
%
\begin{figure}[H]
\haskell{sections/03AAMByExample/02AbstractStateSpace/01AAM.hs}
\caption*{Address and Time Abstraction}
\end{figure}
\noindent
%
An implementation of \h|AAM aam0| for a particular type \p|aam0| must pick two
types, \h|Addr aam0| and \h|Time aam0|, and then implement \h|tzero|, \h|tick|,
\h|alloc|, for those types.
%
The type |aam0| its self will only used as a type level token for selecting
associated types at the type level, and will have a single inhabitant for
selecting type class functions at the value level.
%
(If Haskell had a proper module system this would all be better expressed as a
module interface--containing a type and functions over that type--rather than
an associated type, type class and proxy singleton type.)