### 2014-08-04 tex/sections/03AAMByExample/07Optimizations/02Composition.tex

Heap widening and abstract garbage collection were introduced independently;
what happens when they compose?
%
The result is semantically what you want but exhibits bad behavior
operationally.
%
The naive composition of the optimizations results in each store being garbage
collected independently of each other, which is expensive.
%
What is worse, \textit{they are all the exact same store}, making the extra
effort entirely redundant.

%--%

The solution to this is to either rewrite the semantics to use a widened store
throughout, or to rewrite abstract garbage collection specifically for the
widened state space.
%
Granted, none of these options are terribly difficult in this setting, but
rewriting semantics is a time-consuming and bug-ridden process in general.
%
A more systematic abstraction is clearly needed for these optimizations: one
where optimizations compose seamlessly with both one another and the abstract
semantics.