### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/main.tex

Monadic AAM begins with recognizing that the abstract semantics from
section~\ref{section:AAMByExample:AbstractSemantics} look suspiciously monadic.
%
In this section we will rewrite the exact same function but in monadic style.
%
After the exercise, the monad will be generalized in
section~\ref{section:MonadicAAM:Generalizing} and this generalization will be
exploited in section~\ref{section:MonadicAAM:Optimizations}.

\subsubsection{Nondeterminism}
\input{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/Z.tex}

\subsubsection{State}
\input{sections/04MonadicAAM/00MonadicStyle/01State/Z.tex}