### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/Z.tex

Widening can be recovered by simply propagating the state transformer
containing the heap to the very bottom of the stack.
%
The state space
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/00BeforeSS.hs}
%
induced by the abstract monad
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/01BeforeM.hs}
%
must turn into
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/02AfterSS.hs}
%
the desired state space for heap widening.
%
We show that this state space is \textit{induced purely by a reordering of the
monads} in \h|A_M|, lifting \h|Nondet| up one level and \h|StateT (S d aam)| to
the bottom.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/03AfterM.hs}
%
This permutation is only valid if the resulting monad adheres to the
\h|Semantics| interface, which we justify next.

%--%

To show that \h|A_Widen_M| implements \h|Semantics| we need to invent a
transformer variant of \h|Nondet|, which we call \h|NondetT|.
%
Rather than a pure list of values, \h|NondetT| carries a list of values inside
a monadic action.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/04NondetT.hs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Aside: List Monad Transformer %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph{Aside on List Monad Transformers}
{\it
  It is folklore that the naive implementation of \h|ListT m a| $\equiv$ 
  \h|m [a]| is not a monad.
  %
  Variants of \h|ListT m a| have been proposed which are more akin to monadic
  streams of \p|a| values in the monad \p|m|.
  %
  We are able to interpret \h|m [a]| as both a monad and monad plus given extra
  conditions on the underlying monad, and is to our knowledge the first instance
  where \h|m [a]| is given a valid monadic instance.
}

%%%%%%%%%%%%%%%
%% End Aside %%
%%%%%%%%%%%%%%%

The \h|Monad| instance for \h|NondetT| requires the underlying monad to be a
\h|JoinLattice| \textit{functor}.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/05NondetTMonad.hs}
%
\h|NondetT| also propagates the \h|MonadPlus| interface
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/06NondetTMonadPlus.hs}
%
the \h|MonadState| interfaces
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/07NondetTMonadState.hs}
%
and implements \h|MonadStep|.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/08NondetTMonadStep.hs}

%--%

Given the permuted state space, heap widening is recovered merely by swapping
\h|A_M| with \h|A_Widen_M|.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/09Exec.hs}