### 2014-09-17 writeup/sections/03AAMByExample/04RecoveringConcrete/main.tex
writing final version of things...

The final delta abstraction resulted in a highly factorized state space,
amenable to a wide range of analyses.
%
Before manipulating these axes of parameterization, we first show how to
recover a concrete interpreter.

%--%

Our semantics has two axes of parameterization: \h|AAM| and \h|Delta|.
%
For \h|AAM| we choose integers for abstract time and integers paired with
variable names for abstract addresses.
%
\haskell{sections/03AAMByExample/04RecoveringConcrete/00AAM.hs}

%--%

For \h|Delta| we choose the value type and primitive operations from
section~\ref{section:AAMByExample:Concrete}.
%
\haskell{sections/03AAMByExample/04RecoveringConcrete/01Delta.hs}

%--%

Running the abstract semantics from
section~\ref{section:AAMByExample:AbstractSemantics} with these two
parameterizations--\h|C_AAM| and \h|C_Delta|--recovers the concrete
(collecting) semantics.
%
\haskell{sections/03AAMByExample/04RecoveringConcrete/02Exec.hs}