### 2014-09-18 writeup/old/sections/03AAMByExample/02AbstractStateSpace/02NonDet.tex

Following the AAM methodology, we realize that in a world with finite adresses
but potentially infinite executions, there is a chance we will need to reuse an
existing address.
%
This causes us to introduce a powerset both around the state space and into the
domain of the heap, as a single address may now point to multiple values, and
variable reference may return multiple values.
%
\begin{figure}[H]
\haskell{sections/03AAMByExample/02AbstractStateSpace/02NonDet.hs}
\caption*{Introducing Nondeterminism}
\end{figure}
\noindent
%
As a result of this abstraction, the semantics must consider multiple control
paths where before there was only one.