### 2014-09-18 writeup/old/sections/03AAMByExample/02AbstractStateSpace/main.tex

To arrive at a computable analysis we must abstract the infinite concrete state
space with a finite abstract state space.
%
As we sketch the abstract state space in the AAM methodology we will hint at
corrosponding changes that need to take place in the semantics.
%
The full abstract semantics for the final abstract state space will be given in
section~\ref{section:AAMByExample:AbstractSemantics}.

%--%

\subsubsection{Cutting Recursion}
\input{sections/03AAMByExample/02AbstractStateSpace/00Cut.tex}

\subsubsection{Address and Time Abstraction}
\input{sections/03AAMByExample/02AbstractStateSpace/01AAM.tex}

\subsubsection{Introducing Nondeterminism}
\input{sections/03AAMByExample/02AbstractStateSpace/02NonDet.tex}

\subsubsection{Delta Abstraction}
\input{sections/03AAMByExample/02AbstractStateSpace/03Delta.tex}