### 2014-09-18 writeup/old/sections/03AAMByExample/05Recovering0CFA/main.tex

To recover 0CFA we use variables themselves as abstract addresses and ignore
the time component of the state space.
%
\haskell{sections/03AAMByExample/05Recovering0CFA/00AAM.hs}

%--%

Although the \h|ZCFA_AAM| parameter captures the behavior of 0CFA, we do not
yet have a computable analysis--we need a finite instantiation for the
\h|Delta| parameter.
%
We pick a simple \h|Delta| instance for now, but note that it is not inherent
to 0CFA, it merely serves as \textit{some} abstraction for literals and
primitive operations.
%
\haskell{sections/03AAMByExample/05Recovering0CFA/01Delta.hs}
%
We call this \h|Sym_Delta| because it treats literals purely symbolically,
essentially carrying around only their type information.

%--%

Running the abstract semantics from
section~\ref{section:AAMByExample:AbstractSemantics} with these two
parameterizations--\h|ZCFA_AAM| and \h|Sym_Delta|--recovers a 0CFA analysis.
%
\haskell{sections/03AAMByExample/05Recovering0CFA/02Exec.hs}