### 2014-09-18 writeup/old/sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/Z.tex

Before introducing the nondeterminism monad, we first describe its interface,
\h|MonadPlus|.
%
\h|MonadPlus| abstractly encodes the notion of nondeterminism.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/00MonadPlus.hs}
%
\h|mzero| encodes failure and \h|mplus| encodes nondeterministic choice.

%--%

The nondeterminism monad is built with the list datatype
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/01Nondet.hs}
%
and implements the \h|MonadPlus| interface:
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/02NondetMonadPlus.hs}
%
For example, \h|[1,3] >>= \ x -> [x, x+1]| and \h|[1,3] `mplus` [2,4]| both
evaluate to \h|[1,2,3,4]|.

%--%

The \h|atom| denotation function from
section~\ref{section:03AAMByExample:AbstractSemantics} can be refactored into
the nondeterminism monad by replacing \h|setSingleton| with \h|return|,
\h|eachInSet| with \h|(>>=)|, and \h|setEmpty| with \h|mzero|.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/03Atom.hs}

%--%

\h|call| can be factored similarly.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/04Call.hs}