### 2014-09-18 writeup/old/sections/04MonadicAAM/02RecoveringConcrete/main.tex

To recover concrete semantics we can reuse the concrete parameters \h|C_Delta|
and \h|C_AAM|, and need only design a concrete monad to instantiate the
semantics.

%--%

When implementing concrete semantics in section
\ref{section:AAMByExample:RecoveringConcrete} we were forced to reuse the
\h|Set| type even though the semantics were fully deterministic.
%
Because we have abstracted the semantics over a monad, we are now able to
choose an appropriately precise domain for execution.
%
A more appropriate domain would be the \h|Point| type, where the meaning of
\h|Point a| is the extension of \p|a| with explicit \h|Top| and \h|Bot| values.
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/00Point.hs}

%--%

Point can be shown to be a monad:
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/01PointMonad.hs}
%
and even a monad plus:
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/02PointMonadPlus.hs}
%
and finally a monad step:
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/03PointMonadStep.hs}

%--%

Our final monad stack for recovering concrete semantics will be
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/04ConcreteMonad.hs}
%
and execution of the concrete semantics can be recovered by
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/05Exec.hs}