### 2014-09-18 writeup/old/sections/04MonadicAAM/03RecoveringKCFA/main.tex

We skip the recovery of 0CFA in the monadic semantics and head straight to
kCFA.
%
The only thing that changes from the monadic point of view is the arbiter of
nondeterminism at the bottom of the transformer stack.
%
Rather than \h|Point| we use the \h|Nondet| monad as originally introduced in
section~\ref{section:MonadicAAM:MonadicStyle}.

%--%

kCFA can be recovered with monad stack
%
\haskell{sections/04MonadicAAM/03RecoveringKCFA/00AbstractMonad.hs}
%
and the \h|Delta| and \h|AAM| components described in
section~\ref{section:AAMByExample:RecoveringKCFA}
%
\haskell{sections/04MonadicAAM/03RecoveringKCFA/01Exec.hs}