### 2014-09-29 writeup-old/sections/03AAMByExample/07Optimizations/main.tex
draft is starting to look good, but still not perfect

We now introduce two optimizations to AAM--heap widening and abstract garbage
collection--and show how they integrate into the computational framework
described thus far.

%--%

\subsubsection{Heap Widening}
\input{sections/03AAMByExample/07Optimizations/00Widen.tex}

\subsubsection{Abstract Garbage Collection}
\input{sections/03AAMByExample/07Optimizations/01AbstractGC.tex}

\subsubsection{Composing Heap Widening and Abstract GC}
\input{sections/03AAMByExample/07Optimizations/02Composition.tex}