### 2014-09-29 writeup-old/sections/04MonadicAAM/main.tex

We now present monadic AAM.
%
Monadic AAM is part an extension of the AAM methodology, and part extension of
the AAM toolkit.
%
In methodology, interpreters are first abstracted in the AAM tradition, and
then further factored to be monadic.
%
In toolkit, several monads are provided by our library to recover various
analysis properties in a language agnostic fashion.

\subsection{AAM in Monadic Style}
\label{section:MonadicAAM:00}
\input{sections/04MonadicAAM/00MonadicStyle/main.tex}

\subsection{Generalizing the Monad}
\label{section:MonadicAAM:01}
\input{sections/04MonadicAAM/01Generalizing/main.tex}

\subsection{Recovering a Concrete Interpreter}
\label{section:MonadicAAM:02}
\input{sections/04MonadicAAM/02RecoveringConcrete/main.tex}

\subsection{Recovering kCFA}
\label{section:MonadicAAM:03}
\input{sections/04MonadicAAM/03RecoveringKCFA/main.tex}

\subsection{Optimizations}
\label{section:MonadicAAM:04}
\input{sections/04MonadicAAM/04Optimizations/main.tex}