### 2014-10-25 writeup/snips/Lam-IF-Semantics-Rules.tex
first push on paper content.  focusing on ripping out features piecewise and relating making correct-by-construction arguments along the way.

\begin{alignat*}{1}
  ⟨ a , ρ , [ □ \op e ] ∷ \overline{fr} ⟩ &\steps ⟨ e                         , ρ , [ 𝒜_ρ⟦a⟧ \op □ ] ∷ \overline{fr} ⟩         \\
  ⟨ a , ρ , [ v \op □ ] ∷ \overline{fr} ⟩ &\steps ⟨ \delta(\op , v , 𝒜_ρ⟦a⟧ ) , ρ , \overline{fr} ⟩                            \\
%
  ⟨ a , ρ , [ □                   \ap e ] ∷ \overline{fr} ⟩ &\steps ⟨ e , ρ               , [ 𝒜_ρ⟦a⟧ \ap □ ] ∷ \overline{fr} ⟩ \\
  ⟨ a , ρ , [ ⟨ \lam(x) → e, ρ' ⟩ \ap □ ] ∷ \overline{fr} ⟩ &\steps ⟨ e , ρ'[x ↦ 𝒜_ρ⟦a⟧ ] , \overline{fr} ⟩                    \\
%
  ⟨ a , ρ , [ \ifz{□}{e₁}{e₂} ] ∷ \overline{fr} ⟩ &\steps 
  \begin{cases} 
      ⟨ e₁ , ρ , \overline{fr} ⟩ & \text{if $𝒜_ρ⟦a⟧ = 0$}            \\
      ⟨ e₂ , ρ , \overline{fr} ⟩ & \text{otherwise} 
  \end{cases} \\
%
  ⟨ e₁ \op e₂        , ρ , \overline{fr} ⟩ &\steps ⟨ e₁   , ρ , [ □ \op e₂ ]        ∷ \overline{fr} ⟩               \\
  ⟨ e₁ \ap e₂        , ρ , \overline{fr} ⟩ &\steps ⟨ e₁   , ρ , [ □ \ap e₂ ]        ∷ \overline{fr} ⟩               \\
  ⟨ \ifz{e₁}{e₂}{e₃} , ρ , \overline{fr} ⟩ &\steps ⟨ e₁   , ρ , [ \ifz{□}{e₂}{e₃} ] ∷ \overline{fr} ⟩
\end{alignat*}