### 2014-11-12 writeup/pldi15.tex
added support for raw characters and rearranged macro expansion

%lambdlambdaa any author declaration will be ignored  when using 'plid' option (for double blind review)
\documentclass[pldi,nocopyrightspace]{sigplanconf}

% From PLDI Template {-{

% \usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage{amsmath}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI

% }-}

% Setup {-{

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{stmaryrd}
\usepackage{galois}
\usepackage{float}
\floatstyle{boxed}
\restylefloat{figure}

\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

% This makes newlines not introduce paragraphs.
% Paragraphs must be explicitly marked with \par
% WHAT WIZARDRY IS THIS
\endlinechar=9\relax

% }-}

% Macros {-{

\newcommand\litM{M}
\newcommand\litTime{Time}
\newcommand\litStore{Store}
\newcommand\litVal{Val}

\newcommand{\itop}[1]{\operatorname{\mathit{#1}}}
\newcommand{\ttbfop}[1]{\operatorname{\mathtt{\boldsymbol{\mathbf{#1}}}}}
\newcommand{\ttbfbin}[1]{\mathbin{\mathtt{\mathbf{\boldsymbol{#1}}}}}
\newcommand{\ttop}[1]{\operatorname{\mathtt{#1}}}
\newcommand{\Concrete}[1]{\mathtt{\mathbf{#1}}}
\newcommand{\Abstract}[1]{\widehat{\mathtt{\mathbf{#1}}}}

% }-}

\begin{document}

% These are supposed to make align* not have breaks around it...
\setlength{\abovedisplayskip}{0em}
\setlength{\belowdisplayskip}{0em}
\setlength{\abovedisplayshortskip}{0em}
\setlength{\belowdisplayshortskip}{0em}

\title{Modular Metatheory for Abstract Interpreters}
\maketitle

% Abstract {-{
\begin{abstract}

The design and implementation of static analyzers have becoming increasingly systematic.
In fact, design and implementation have remained seemingly on the verge of full mechanization for several years.
A stumbling block in full mechanization has been the ad hoc nature of soundness proofs accompanying each analyzer.
While design and implementation is largely systematic, 
  soundness proofs can change significantly with (apparently) minor changes to the semantics and analyzers themselves.
We finally reconcile the systematic construction of static analyzers with their proofs of soundness 
  via a mechanistic Galois-connection-based metatheory for static analyzers.

\end{abstract}
% }-}

\input{tmp/autogen/pldi15.markdown.tex}

\end{document}