### 2015-03-23 writeup/oopsla2015.tex
removed authors and categories etc.

\documentclass[nocopyrightspace,preprint,10pt]{sigplanconf}

% OOPSLA Setup {-{

% % \usepackage{SIunits}            % typset units correctly
% \usepackage{courier}            % standard fixed width font
% \usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
% \usepackage{url}                  % format URLs
% \usepackage{listings}          % format code
% \usepackage{enumitem}      % adjust spacing in enums
% \usepackage{amsmath}
% \usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% % known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
% \newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI

% }-}

% Setup {-{

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{stmaryrd}
\usepackage{galois}
\usepackage{float}
\usepackage{latexsym}
\usepackage{balance}
\floatstyle{boxed}
\restylefloat{figure}
\usepackage[font=small,skip=0pt]{caption}

\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}

% This makes newlines not introduce paragraphs.
% Paragraphs must be explicitly marked with \par
% WHAT WIZARDRY IS THIS
\endlinechar=9\relax

% }-}

% Macros {-{

\newcommand\litM{M}
\newcommand\litTime{Time}
\newcommand\litStore{Store}
\newcommand\litVal{Val}
\newcommand\litKAddr{KAddr}
\newcommand\litKStore{KStore}
\newcommand\litEnv{Env}
\newcommand\litClo{Clo}
\newcommand\litelse{else}

\newcommand{\itop}[1]{\operatorname{\mathit{#1}}}
\newcommand{\ttop}[1]{\operatorname{\mathtt{#1}}}
\newcommand{\ttbfop}[1]{\ttop{\boldsymbol{\mathbf{#1}}}}
\newcommand{\ttbfbin}[1]{\mathbin{\mathtt{\mathbf{\boldsymbol{#1}}}}}
\newcommand{\keyword}[1]{\underline{\ttbfop{#1}}}
\newcommand{\keywordop}[1]{\ttbfbin{#1}}
\newcommand{\Concrete}[1]{\mathtt{\mathbf{#1}}}
\newcommand{\Abstract}[1]{\widehat{\mathtt{\mathbf{#1}}}}

% }-}

\begin{document}

% OOPSLA Prelude {-{

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% \conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
% \copyrightyear{20yy} 
% \copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% -- I'm not sure which one of these we use (or none?)... -DCD
%
% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{Preprint}             % These are ignored unless
% \preprintfooter{preprint footer}   % 'preprint' option specified.

% }-}

% These are supposed to make align* not have breaks around it...
\setlength{\abovedisplayskip}{0em}
\setlength{\belowdisplayskip}{0em}
\setlength{\abovedisplayshortskip}{0em}
\setlength{\belowdisplayshortskip}{0em}

\title{Galois Transformers and Modular Abstract Interpreters}
\subtitle{Reusable Metatheory for Program Analysis}
% \authorinfo{David Darais}{University of Maryland, USA}{darais@cs.umd.edu}
% \authorinfo{Matthew Might}{University of Utah, USA}{might@cs.utah.edu}
% \authorinfo{David Van Horn}{University of Maryland, USA}{dvanhorn@cs.umd.edu}
\authorinfo{}{}{}
\maketitle

% Abstract {-{
\begin{abstract}

The design and implementation of static analyzers have becoming increasingly
systematic.  In fact, for large classes of analyzers, design and implementation
have remained seemingly (and now stubbornly) on the verge of full mechanization
for several years.  A stumbling block in full mechanization has been the
\emph{ad hoc} nature of soundness proofs accompanying each analyzer. While
design and implementation is largely systematic, soundness proofs can change
significantly with seemingly minor changes to the semantics or analyzers.  An
achievement of this work is to systematize, parameterize and modularize the
proofs of soundness, so as to make them composable across analytic properties.

\par

We solve the problem of systematically constructing static analyzers by
introducing \emph{Galois transformers}: monad transformers that
transports Galois connection properties.  In concert with a monadic
interpreter, we define a library of monad transformers that
implement building blocks for classic analysis parameters like context-, path-,
and heap-(in-)sensitivity. Moreover, these can be composed together 
% to realize a program analysis monad, 
\emph{independent of the language being analyzed}.

\par

Significantly, a Galois transformer can be proved sound once and for all,
making it a reusable analysis component.  As new analysis features and
abstractions are developed and mixed in, soundness proofs need not be
reconstructed, as the composition of a monad transformer stack is sound by
virtue of its constituents.  Galois transformers provide a viable foundation
for reusable and composable metatheory for program analysis.

\par

Finally, these Galois transformers shift the level of abstraction in analysis
design and implementation to a level where non-specialists have the ability to
synthesize sound analyzers over a number of parameters.

\end{abstract}
% }-}

% \category{CR-number}{subcategory}{third-level}
% \terms term1, term2
% \keywords keyword1, keyword2

\input{tmp/autogen/paper.markdown.tex}

% \appendix
% \section{Appendix Title}
% 
% Appendix body.
% 
% \acks
% 
% Acknowledgments body.

\balance
\bibliographystyle{abbrvnat}
\bibliography{dvanhorn,davdar}

\end{document}